### 2017-03-27 thesis/results/results.tex
edited in Mik's comments and spell-checked everything

\begin{figure}[tb!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the heirarchical clustering of the genes is shown on the left of the heatmap.
	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
	\label{fig:crmetaheat}
\end{figure}

@@

These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the orignial data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).

@@

Both the obesity associated genetic signatures identfied from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
To invevstigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.

Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity associated genetic signature from the original CR data.
\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
% Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma} normalised CR data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.

@@

\begin{table}[htpb]
	\centering
	\caption[Summary of the number of \acrshort{deg}s identified using different unadjusted and \acrshort{fdr}-adjusted p-value threshold in different versions of the \gls{rma} normalised CR data set]{Summary of the number of \glspl{deg} identified using different unadjusted and \gls{fdr}-adjusted p-value thresholds in different versions of the \gls{rma} normalised CR data set}
	\label{tab:ob_deg_summary}
	\begin{tabular}{lcccc}
		& \multicolumn{4}{c}{Number of \glspl{deg} identified}\\
		\cmidrule(r){2-5}
		& \multicolumn{2}{c}{Unadjusted p-value} & \multicolumn{2}{c}{\gls{fdr}-adjusted p-value}\\
		\cmidrule(r){2-3} \cmidrule(r){4-5}
		& 0.01 & 0.05 & 0.01 & 0.05\\
		\hline
		\hline
		\rule{0pt}{2.25ex}CR data       & 1781 & 5278 & 0 & 9 \\
		Residual CR data                & 1104 & 4371 & 0 & 0 \\
		Caucasian-only CR data          & 2129 & 6029 & 0 & 0 \\
		Residual Caucasion-only CR data & 1558 & 5427 & 0 & 0 \\
		\hline
		\hline
	\end{tabular}
\end{table}

@@

With the Caucasian-only CR data set, obesity associated genetic signatures were identified as described above.
% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
2129 and 1558 gene probes were significantly differentially expressed (unadjuste p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
As before, the most significant 799 gene probes were selected from these gene probes.
There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).

@@

Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.

@@

When the pathway enrichment analysis was carried out with the \gls{kegg} pathway database, only the ``ABC transporter'' was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
With the Reactome database, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
With the \gls{go} database, there were 22 pathways significantly enriched in \gls{blca}, 17 pathways in \gls{cesc}, 3 pathways in \gls{kirp}, 21 pathways in \gls{read}, 10 pathways in \gls{skcm}, 14 pathways in \gls{ucec}, and no pathways were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).

Though there were many pathways identified as significantly enriched with the \gls{go} database in each of the cancer types, there were no pathways that were common across all of the different cancer types.
Furthermore, these pathways were not similar in terms of the biological activities involved.
For example, pathways enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but the pathways enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three pathway databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
All of these results indicated that the pathways that drive tumour progression were unique to each cancer type, and no one pathway was associated with obesity and tumour biology across multiple cancer types.

Since there were no common pathway enriched in the \gls{icgc} cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined into a single data set to see if there were any pathways enriched in the combined data set.
The combined data set was generated in two ways.
The first combined data set was created by voom normalising each cancer data set separately, combine all of the normalised cancer data sets into one, then applied batch correction to the data set (\cref{sub:batch_correction}).
The second combined data set was created by combining the cancer data sets into one big data set first, then the combined data was voom normalised, and finally batch correction was applied to obtain the final data set.
Each of the combined data sets were analysed for enriched pathways, but there were no pathways significantly enriched in either data sets.
These results suggested that there were no common biological pathways associated with obesity in the \gls{icgc} cancer data sets.

>>>>

\begin{figure}[tb!]
	\centering
	\includegraphics[page=3,width=0.8\linewidth]{results1/creighton_mg_heatmap1}
	\caption[Obesity metagene from \citet{Creighton2012} study and sample gene expression in CR data] {Heatmap showing the obesity metagene from the \citet{Creighton2012} study with the sample gene expressions of the obesity associated genes in the CR data.
	Level of expression is represented in the top right histogram, where low and high gene expression are colour-coded with blue and red, respectively.
	Each row of the heatmap represents a gene from the obesity associated genetic signature, and each column of the heatmap represents a sample from the CR data.
	The obesity associated metagene scores of the samples are shown as a separate row at top of the heatmap, and the tree diagram of the hierarchical clustering of the genes is shown on the left of the heatmap.
	For clarity, the sign of the metagene scores were reversed in order to match the results from the \citet{Creighton2012} study.}
	\label{fig:crmetaheat}
\end{figure}

@@

These results showed that the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study was not generalisable to other cancer data sets, similar to the obesity signature identified by Creighton \textit{et al.}.
This meant that both the Creighton \textit{et al.} and Fuentes-Mattei \textit{et al.} obesity metagenes may have been too specific to the original data set in which the signatures were identified in.
Furthermore, there was a possibility that these obesity associated metagenes were not related to obesity, but associated with a different clinical variable that may be closely related to \gls{bmi} (\cref{sec:creighton_obesity_metagene_new}).

@@

Both the obesity associated genetic signatures identified from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies were able to capture the overall gene expression patterns  of the genetic signatures of the samples, but did not associate with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets.
One possible reason for this result could be that the obesity associated genetic signatures from the \citet{Creighton2012} and \citet{Fuentes-Mattei2014} studies may not have been truly associated with patient \gls{bmi}, but with another clinical variable.
To investigate this possibility, novel obesity associated genetic signatures were identified in the CR data after controlling for all the clinical variables in the data set.
The FM data set was not used to get the obesity associated genetic signatures, as no \gls{bmi} information was available for the patients in the FM data.

Firstly, an attempt was made to replicate the Creighton \textit{et al.} obesity associated genetic signature from the original CR data.
\citet{Creighton2012} originally found their obesity associated genetic signature by doing a gene expression analysis between the breast tumour samples from the obese and non-obese patients, and from this list, Creighton \textit{et al.} selected the 799 statistically significant \glspl{deg} (p \textless{} 0.01) with log$_2$-fold change greater than 1.2.
Therefore, gene expression analysis  between the samples from obese and non-obese patients in the \gls{rma} normalised CR data was carried out (described in \cref{sec:gene_expression_analysis}).
% Therefore, \glspl{deg}  were identified between the samples from obese and non-obese patients in the \gls{rma} normalised CR data, as described in \cref{sec:gene_expression_analysis}.
Without adjusting the p-value to account for multiple hypothesis testing, 5278 gene probes and 1781 gene probes were significant at p \textless{} 0.05 and p \textless{} 0.01, respectively (\cref{tab:ob_deg_summary}).
After adjustments were made for multiple hypothesis testing, there were only 9 gene probes significant at p \textless{} 0.05 and no gene was significant at p \textless{} 0.01, which suggest that the majority of the 799 probe sets reported by Creighton \textit{et al.} may actually be false positives.

@@

\begin{table}[htpb]
	\centering
	\caption[Summary of the number of \acrshort{deg}s identified using different unadjusted and \acrshort{fdr}-adjusted p-value threshold in different versions of the \gls{rma} normalised CR data set]{Summary of the number of \glspl{deg} identified using different unadjusted and \gls{fdr}-adjusted p-value thresholds in different versions of the \gls{rma} normalised CR data set}
	\label{tab:ob_deg_summary}
	\begin{tabular}{lcccc}
		& \multicolumn{4}{c}{Number of \glspl{deg} identified}\\
		\cmidrule(r){2-5}
		& \multicolumn{2}{c}{Unadjusted p-value} & \multicolumn{2}{c}{\gls{fdr}-adjusted p-value}\\
		\cmidrule(r){2-3} \cmidrule(r){4-5}
		& 0.01 & 0.05 & 0.01 & 0.05\\
		\hline
		\hline
		\rule{0pt}{2.25ex}CR data       & 1781 & 5278 & 0 & 9 \\
		Residual CR data                & 1104 & 4371 & 0 & 0 \\
		Caucasian-only CR data          & 2129 & 6029 & 0 & 0 \\
		Residual Caucasian-only CR data & 1558 & 5427 & 0 & 0 \\
		\hline
		\hline
	\end{tabular}
\end{table}

@@

With the Caucasian-only CR data set, obesity associated genetic signatures were identified as described above.
% ; first in the unadjusted data, then in the data with clinical variables adjusted (except ethnicity, as it has already been controlled for by considering only the Caucasian samples), and lastly the common genes between each of these two signatures with the original obesity associated genetic signature were identified.
2129 and 1558 gene probes were significantly differentially expressed (unadjusted p-value \textless{} 0.01) in the unadjusted and clinical variable-adjusted Caucasian patient data, respectively (\cref{tab:ob_deg_summary}).
As before, the most significant 799 gene probes were selected from these gene probes.
There were 148 and 92 common gene probes with the original Creighton \textit{et al.} obesity associated genetic signatures and unadjusted or clinical variable-adjusted Caucasian patient data set, respectively.
Again, Venn diagram was used to summarise the genes identified from the Caucasian patient data set (\cref{fig:venn2}).

@@

Since all of these obesity associated genetic signatures were derived based on the discrete values of the patient \gls{bmi} status, an obesity associated genetic signature was searched for using the correlation of the gene probe expression with the patient \gls{bmi} value (continuous variable) in the CR data with all of the samples included, but no significant genes were identified (\cref{sec:obesity_associated_genetic_signature_using_sample_bmi_values}).
All of these obesity associated genetic signatures (eight in total) were checked to see whether these signatures had significant association with patient \gls{bmi} or \gls{bmi} status in the \gls{nzbc} and \gls{icgc} cancer data sets (\cref{sub:_novel_obesity_associated_signatures_and_sample_bmi}).
For simplicity, the abbreviations in \cref{tab:mg_abbrev} will be used to refer to the appropriate genetic signatures.

@@

When the pathway enrichment analysis was carried out with the \gls{kegg} pathway database, only the ``ABC transporter'' pathway was significantly enriched (\gls{fdr}-adjusted p-value \textless{} 0.05) in the \gls{cesc} data, and no other pathways were enriched in any of the other \gls{icgc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
With the Reactome database, ``Phosphatase bond hydrolysis by NUDT proteins'' (in \gls{blca}) and ``Mitochondrial ABC transporters'' (in \gls{cesc}) were significantly enriched (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
With the \gls{go} database, there were 22 \gls{go} terms significantly enriched in \gls{blca}, 17 terms in \gls{cesc}, 3 terms in \gls{kirp}, 21 terms in \gls{read}, 10 terms in \gls{skcm}, 14 terms in \gls{ucec}, and no terms were significant in the \gls{coad} and \gls{lihc} cancer types (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).

Though there were many \gls{go} terms identified as significantly enriched with the \gls{go} database in each of the cancer types, there were no terms that were common across all of the different cancer types.
Furthermore, these terms were not similar in terms of the biological activities involved.
For example, \gls{go} terms enriched in the \gls{blca} data suggested possible activation of the \gls{pi3k} pathway, but those enriched in the \gls{read} data were mainly involved in \acrshort{rna} processing (\cref{sec:pathways_significant_in_each_of_the_cancer_types}).
One interesting result from this enrichment analysis was that the ``ABC transport'' pathway was identified as significantly enriched in the \gls{cesc} data set by all three databases, suggesting that the ``ABC transport'' pathway may be a core component in the \gls{cesc} tumour biology.
All of these results indicated that the biological processes that drive tumour progression were unique to each cancer type, and no one pathway was associated with obesity and tumour biology across multiple cancer types.

Since there were no common pathway enriched in the \gls{icgc} cancer types when analysed individually, all of the \gls{icgc} cancer data sets were combined into a single data set to see any enriched pathways could be identified in a collective analysis.
The combined data set was generated in two ways.
The first combined data set was created by normalising (via voom) each cancer data set separately, combine all of the normalised cancer data sets into one, then applying batch correction to the data set (\cref{sub:batch_correction}).
The second combined data set was created by combining the cancer data sets into one big data set first, then the combined data was normalised, and finally batch correction was applied to obtain the final data set.
Each of the combined data sets were analysed for enriched pathways, but there were no pathways significantly enriched in either data sets.
These results suggested that there were no common biological pathways associated with obesity in the \gls{icgc} cancer data sets.

### 2017-03-27 thesis/results/results2.tex
edited in Mik's comments and spell-checked everything

In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
However, the analyses so far have used \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no significant difference in the ranking approaches used.

Scatter plots were plotted and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
The GT pathway metagenes were generated by using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
From these results, it was clear the normalisation methods used on the data set had the most significant effect on the resulting metagenes, rather than the normalisation methods used to generate the transformation matrices (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
This meant that all of the data sets had to be normalised with a single normalisation method, so the \gls{rma} was chosen as the normalisation method to be used as this method was more reliable than the \gls{mas} method \citep{Irizarry2003}.

Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
Taken together, it was decided that the GT data set was batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

@@

In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
Though \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.

FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
This may have been due to the lack of consistency of the Akt pathway signature in the different data sets (\cref{tab:svd_vs_tm_path}).
Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from the other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.

@@

The models constructed from the patient \gls{bmi}, \gls{bmi} status and/or GT pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for the CR metagenes and the FM metagene, respectively.
It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in \gls{nzbc} data set, as majority of the patients in the \gls{nzbc} data set were \gls{pr}$^+$ (\cref{tab:clin_summary}).
The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of the FM metagene with the Myc pathway by \citet{Fuentes-Mattei2014}.
\textit{MYC} is an oncogene that encodes the Myc transcription factor that, when overexpressed, disrupts various pathways that directly promote tumour growth and proliferation, and has a central role in many of the cancer hallmarks \citep{Coller2000,Hanahan2000}.
``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason why the Myc pathway metagene showed up as a significant variable in the linear model.

@@

Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative assocation of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.

@@

Clearly, the stepwise linear models were able to predict the obesity associated metagenes better than the models created in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
This could be due to the fact that many of the stepwise linear models included many pathways that had a strong negative correlation with the obesity metagenes (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways; \cref{fig:gatza_allmeta}).
These pathway metagenes may have been better predictors of the obesity metagenes than the six pathways that were ``consistent'' across the data sets, and therefore generated better predictions than those linear models.
With that said, some of the predicted obesity metagenes from the stepwise linear models were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
Again, this could have been due to the model being overfitted in the \gls{nzbc} data, and thus made an incorrect prediction about the obesity metagenes in the CR data.

@@

Taken together, these results suggested that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
However, there were no single common pathway that was present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.

>>>>

In the \citet{Gatza2010a} study, their data comprised of samples from different studies which were aggregated into a single data set that was \gls{mas} normalised, and the metagene scores were transformed to a scale between 0 to 1 using a probit function.
However, the analyses so far have used the \gls{rma} normalisation method and ranked based on the number of samples present in the data (fractional ranking; \cref{sub:ranking_of_the_metagene_scores}).
To decide which normalisation methods or ranking approaches were suitable for the analysis, the different methods were compared in the GT data (\cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures,sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
Results shown in \cref{sec:ranking_method_for_the_pathway_associated_genetic_signatures} clarified that there was no substantive difference in the ranking approaches used.

Scatter plots were generated and the correlations were calculated for each of the GT pathway metagenes derived from either \gls{rma} or \gls{mas} normalised GT data (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
The GT pathway metagenes were generated using the transformation matrices that were derived from either the \gls{rma} or \gls{mas} normalised GT data.
From these results, it was clear the normalisation methods used on the data set had the most significant effect on the resulting metagenes, rather than the normalisation methods used to generate the transformation matrices (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
This meant that all of the data sets had to be normalised with a single normalisation method, so \gls{rma} was chosen as the normalisation method to be used as this method is regarded as being more reliable than the \gls{mas} method \citep{Irizarry2003}.

Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
Taken together, it was decided that the GT data set should be batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

@@

In \cref{fig:gatza_allmeta}, all of the obesity associated metagenes from the CR data set cluster together into a group, but none of the pathway associated metagenes had  a strong positive correlation with the obesity metagenes.
Though the  \gls{her2} and p63 pathway metagenes were grouped next to the cluster of obesity metagenes, the similarity was at most about 0.6  (Pearson correlation; \cref{fig:gatza_allmeta}).
With that said, these obesity associated metagenes seem to have a strong negative correlation with some of the pathway associated genetic signatures, such as the Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways, and this strong negative correlation was observed in the other data sets as well (\cref{sec:correlation_of_the_pathway_and_obesity_metagenes_in_other_cancer_data_sets}).
In fact,  high Or metagene scores related to lower levels of gene expression for the genes in the Or signature (\cref{fig:crmetaheat}; \cref{sec:creighton_obesity_metagene}), and therefore it could be possible that many of the genes in the Or signature (and the other obesity associated signatures from the CR data) may be positively correlated with the genes in these pathway signatures.
This suggests that the obesity associated genetic signatures from the CR data may be related to these pathway signatures, rather than the obesity phenotype of the patients.

The FM obesity associated metagene did not cluster together with the obesity metagenes from the CR data set, nor with any of the other pathway metagenes.
The fact that the FM metagene did not cluster with any of the CR metagenes showed that the signals captures by the FM metagene was different from those metagenes found in the CR data set.
Furthermore, the FM metagene did not cluster with the Akt pathway in the heatmap, even though \citet{Fuentes-Mattei2014} provided clear evidence that their metagene specifically activated the Akt/\gls{mtor} signalling pathway in a mouse model.
This may have been due to the lack of consistency of the Akt pathway signature in the different data sets (\cref{tab:svd_vs_tm_path}).
Another likely reason could be that FM genetic signature was not comprised only of the genes from the Akt pathway, but genes from the other pathways as well, and so the metagene did not cluster together with the Akt pathway signature in the heatmap.

@@

The models constructed from the patient \gls{bmi}, \gls{bmi} status and/or GT pathway metagene scores confirmed that none of the pathways were significant, except for the \gls{pr} and Myc pathway metagene scores in the models for the CR metagenes and the FM metagene, respectively.
It was not surprising that the \gls{pr} pathway metagene showed significance in the models generated in \gls{nzbc} data set, as majority of the patients in the \gls{nzbc} data set were \gls{pr}$^+$ (\cref{tab:clin_summary}).
The significance of the Myc metagene in the model for FM metagene was difficult to interpret, as there was no reported association of the FM metagene with the Myc pathway by \citet{Fuentes-Mattei2014}.
\textit{MYC} is an oncogene that encodes the Myc transcription factor that, when overexpressed, disrupts various pathways that directly promote tumour growth and proliferation, and has a central role in many of the cancer hallmarks \citep{Coller2000,Hanahan2000}.
``Sustained cell proliferation'' was one of the cancer hallmarks that \citet{Fuentes-Mattei2014} had shown to be related to their obesity associated genetic signature, which could be the reason that the Myc pathway metagene showed up as a significant variable in the linear model.

@@

Another thing to note from \cref{fig:predict_cr_cris} was the apparent negative association of the predicted metagenes with the original metagenes in the \gls{bmi}-only, the \gls{bmi} status-only and the \gls{bmi} and \gls{bmi} status models.
One possible explanation for this was that these \gls{bmi}-based models were overfitted in the \gls{nzbc} data set, and was not generalisable in the CR data set, and thus produced negative correlations.
In fact, it was evident from the metagene analyses in \cref{cha:obesity_genetic_signatures_and_cancer} that these obesity associated genetic signatures were not generalisable with the patient \gls{bmi} in different data sets, and overfitting of the model in the \gls{nzbc} data was the likely cause of this negative correlation.

@@

Clearly, the stepwise linear models were able to predict the obesity associated metagenes better than the models created in \cref{sub:linear_model_prediction_in_the_nzbc_and_cr_data_sets}.
This could be due to the fact that many of the stepwise linear models included many pathways that had a strong negative correlation with the obesity metagenes (Akt, \gls{er}, \gls{egfr}, \gls{pr}, p53, Src and \gls{tgfb} pathways; \cref{fig:gatza_allmeta}).
These pathway metagenes may have been better predictors of the obesity metagenes than the six pathways that were ``consistent'' across the data sets, and therefore generated better predictions than those linear models.
With that said, some of the predicted obesity metagenes from the stepwise linear models were negatively correlated with the original metagene values in the CR data (\cref{fig:stepwise_cr}).
Again, this could have been due to the model being overfit in the \gls{nzbc} data, and thus making an incorrect prediction about the obesity metagenes in the CR data.

@@

Taken together, these results suggest that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.
